---
title: "Model Comparison"
output: html_notebook
---

## libraries

```{r}
library(pROC)
library(epiR)
```

## load roc models

```{r}
lassoRoc <- readRDS("smallLassoRoc.rds")
rfRoc <- readRDS("rfROC.rds")
gbmRoc <- readRDS("gbmROC.rds")
test <- read.csv("testImputedData4.csv", stringsAsFactors = T)
```

## ROC curves

```{r}
graph <- plot(lassoRoc, col = 1, add = F, legacy.axes = T)
graph <- plot(rfRoc, col = 2, add = T)
graph <- plot(gbmRoc, col = 3, add = T)
legend(0.2, 0.5, lty = 1:3, col = 1:3, legend = c("glmL1", "rf", "gbm"), bty = "n" )
```

## comparison of models by DeLong's test

```{r}
roc.test(lassoRoc,gbmRoc)
roc.test(lassoRoc, rfRoc)
roc.test(gbmRoc, rfRoc)
```
By comparison of AUC's both ensembles are statistically better than the L1regression model.  All of the models have very large AUC's

## model comparison by chi square of predictions

```{r}
LassoMtx <- readRDS("smallLassoMatrix.rds")
rfMtX <- readRDS("rfConfusMTX.rds")
gbmMtx <- readRDS("gbmConfusMTX.rds")
```

```{r}
LassoLowRisk <- LassoMtx$table[1,]
LassoHighRisk <- LassoMtx$table[2,]
rfLowRisk <- rfMtX$table[1,]
rfHighRisk <- rfMtX$table[2,]
gbmLowRisk <- gbmMtx$table[1,]
gbmHighRisk <- gbmMtx$table[2,]
NoInfo <- summary(factor(test$response))
```

## construct test matrices

```{r}
LowRisk <- rbind(NoInfo, LassoLowRisk, rfLowRisk, gbmLowRisk)
HighRisk <- rbind(NoInfo,LassoHighRisk, rfHighRisk, gbmHighRisk)

LRMTX <- matrix(data = LowRisk, nrow = 4, ncol = 2, byrow = F, dimnames = list(c("No Info", "L1REgression", "random forest", "generalized boosted"), c("no", "yes")))

HRMTX <- matrix(data = HighRisk, nrow = 4, ncol = 2, byrow = F, dimnames = list(c("No Info", "L1REgression", "random forest", "generalized boosted"), c("no", "yes")))
```

```{r}
LRMTX
HRMTX
```


## chisquare tests
### lowrisk
```{r}
chisq.test(LRMTX)
chisq.test(LRMTX[-1,])


```

Using low risk predictions for the test set, all 3 models are statistically different from the no information model demostrating utility; the predictions made by the 3 models are not statistically different.

```{r}
LowRiskTable <- as.data.frame(LRMTX)
n <- apply(LowRiskTable, 1, sum)
LowRiskTable$N <- n
LowRiskTable$FNPerCent <- 100*(LowRiskTable$yes/LowRiskTable$N)
LowRiskTable$LowRiskPerCent <- LowRiskTable$N/129512

write.csv(LowRiskTable, file = "SummaryLowRiskTable.csv")

```

### confusion matrix
```{r}
cm <- readRDS("gbmConfusMTX.rds")
CM <- cm$table
rval<- epi.tests(CM, conf.level = 0.95)
rval
```
```{r}
cmRF <- rfMtX$table
rvalRF<- epi.tests(cmRF, conf.level = 0.95)
rvalRF
```
```{r}
cmLasso <- LassoMtx$table
rvalLasso<- epi.tests(cmLasso, conf.level = 0.95)
rvalLasso
```



### high risk predictions

```{r}
chisq.test(HRMTX)
chisq.test(HRMTX[-1,])
```
The 3 models are not statistically different in predicting high risk patients


```{r}
HighRiskTable <- as.data.frame(HRMTX)
n <- apply(HighRiskTable, 1, sum)
HighRiskTable$N <- n
HighRiskTable$TPPerCent <- 100*(HighRiskTable$yes/HighRiskTable$N)
HighRiskTable$HighRiskPerCent <- HighRiskTable$N/129512
HighRiskTable

```

write.csv(HighRiskTable, file = "SummaryHighRiskTable.csv")
```