---
title: "Ensemble models using h20"
output: html_notebook
---

## libraries
```{r}
library(h2o)
library(caret)
library(pROC)
```

## data 
```{r}
train <- read.csv("TrainSelectedFeatures.csv", stringsAsFactors = T)[,-1]
test <- read.csv("testImputedData4.csv", stringsAsFactors = T)[,-1]


```

## initiate h2o and upload datasets

```{r}
h2o.init(nthreads = 3)
train.h2o <- as.h2o(train)
test.h2o  <- as.h2o(test)
```


## define y and x

```{r}
y <- "response"
x <- setdiff(names(train.h2o), y)
```

##random forest model

### tune model using grid search

```{r}
rfTuned <- h2o.grid("randomForest",
              search_criteria = list(
                strategy = "RandomDiscrete",
                max_models = 8,
                max_runtime_secs = 600),
              hyper_params = list(
                ntrees = seq(100, 1000, 100),
                max_depth = seq(6,20,1)),
              grid_id = "Random_8",
              x = x,
              y = y,
              training_frame = train.h2o,
              nfolds = 5
              )              
```
```{r}
summary(rfTuned)

```
## generate best rf model

```{r}
rfBest <- h2o.randomForest(y = y, x = x, 
                           training_frame = train.h2o,
                           nfolds = 5,
                           ntrees = 62, 
                           max_depth = 10)

```

## cross validated error
```{r}
h2o.performance(rfBest, test.h2o)
```

## predictions with rfBest

```{r}
rfBestPred <- h2o.predict(rfBest, newdata = test.h2o)
```
### return predictions to local site and create roc models

```{r}
rfPredictions <- as.data.frame(rfBestPred)
```


```{r}
rf.roc <- roc(response = test$response, predictor = rfPredictions[,2], auc = T)
auc(rf.roc)
```


```{r}
coords(rf.roc, x = "best", best.method = "youden", transpose = T)
```

## confusion marix at threshold
```{r}
rfResponse <- as.factor(ifelse(rfPredictions[,2] >= .9989309, "No", "Yes"))
rfmatrix <- confusionMatrix(rfResponse, test$response)
rfmatrix
```
The false negative rate with no information is .15% At threshold for the model, The false negative rate for low risk patients is .008%%;  10.2% of patients would qualify for treatment. The risk for DVT in the high risk group is 1.20%.


### save rf output

```{r}
saveRDS(rf.roc, file = "rfROC.rds")
saveRDS(rfPredictions,  file = "rfPreds.rds")
saveRDS(rfmatrix, file = "rfConfusMTX.rds")
rfModel2 <- h2o.saveModel(rfBest, path = "C:/Users/Stephen Aronoff/Dropbox/DVT study/analysis of cleaned data")
```


## gbm model

```{r}
gbmTuned <- h2o.grid("gbm",
              search_criteria = list(
                strategy = "RandomDiscrete",
                max_models = 8,
                max_runtime_secs = 900),
              hyper_params = list(
                ntrees = seq(50, 1000, 50),
                max_depth = seq(6,20,1),
                min_rows = 500,
                seed = 123,
                learn_rate = 10^seq(-5,0, 1)),
              
              grid_id = "Random_8gbm",
              x = x,
              y = y,
              training_frame = train.h2o,
              nfolds = 5,
              distribution = "bernoulli",
              balance_classes = T
              )


```


```{r}
summary(gbmTuned)
```

```{r}
gbmBest <- h2o.gbm(x = x, y = y, training_frame = train.h2o,
                   nfolds = 5,
                   learn_rate = 0.1,
                   max_depth = 7, 
                   min_rows = 500,
                   ntrees = 47,
                   seed = 123,
                   distribution = "bernoulli", 
                   balance_classes = T)
``` 

```{r}
summary(gbmBest)

```

```{r}
gbmpred <- h2o.predict(gbmBest, newdata = test.h2o)
gbmPredictions <- as.data.frame(gbmpred)
```

```{r}
gbm.roc <- roc(response = test$response, predictor = gbmPredictions[,2], auc = T)
auc(gbm.roc)
```


```{r}
coords(gbm.roc, x = "best", best.method = "youden", transpose = T)
```

```{r}
gbmResponse <- as.factor(ifelse(gbmPredictions[,2] >= .9989878, "No", "Yes"))
gbmMatrix <- confusionMatrix(gbmResponse, test$response)
gbmMatrix
```
The false negative rate with no information is .15% At threshold for the model, The false negative rate for low risk patients is .008%;  8.0% of patients would qualify for treatment. The risk for DVT in the high risk group is 1.47%.


## save models and rocs

```{r}
saveRDS(gbm.roc, file = "gbmROC.rds")
saveRDS(gbmPredictions,  file = "gbmPreds.rds")
saveRDS(gbmMatrix, file = "gbmConfusMTX.rds")
gbmBest <- h2o.saveModel(gbmBest, path = "C:/Users/Stephen Aronoff/Dropbox/DVT study/analysis of cleaned data")
```



